---
title: "Tidyverse Exercise"
author: ""
date: "22/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidyverse

- Load packages
  - we need tidyverse

```{r}

```

## Section 1: Data description

### read csv data

We will read the file "data/covid_us_county.csv.gz"

```{r}
data_covid <- NA

```

### Simple discription of the table

Now the data is in R workspace. Do the following


### Count the number of variabes and rows

```{r}

```

### Print out the first 6 rows of the data.frame

```{r}

```


### How many states exists?

(Hint: use `count()`)


```{r}

```


## Data wrangling, Part 1

### Erase unnecessary rows

First remove the non-country entries

```{r}

```


### Create a subset dataset

Find the latest date of the data, then subset the data with only the newest date for each country

```{r}

```

### Max cases/deaths

Which county has the highest number of cases/deaths? What is the number?

```{r}

```


## Data wrangling, Part 2


### First recorded cases/deaths (by county)

Find the 10 counties with the earliest recorded cases/deaths.

```{r}

```

### Aggregate the number of death

For each state, calculate the total number of cases/deaths for each day (Hint: grouping and summarize)


```{r}

```


### First recorded cases/deaths (by state)

Find the earliest day when a state recorded a case. Arrange by the date in descending order


## Visualizing (Optional)

Now let's visualize the outputs using ggplot

### Time series plot of total cases/death in the US

```{r}

```


### Time series plot of total cases/deaths by the state

```{r}

```


### Daily increase trend

How's the trend of the daily case increases. Which state look bad now?

```{r}

```
