---
title: "Tidyverse Exercise"
author: "Philipp Umek"
date: "22/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidyverse

- Load packages
  - we need tidyverse

```{r}
library(tidyverse)
```

## Section 1: Data description

### read csv data

We will read the file "data/covid_us_county.csv.gz"

```{r}
data_covid <- read_csv("data/covid_us_county.csv.gz")

```

### Simple discription of the table

Now the data is in R workspace. Do the following


### Count the number of variabes and rows

```{r}
dim(data_covid)
```

### Print out the first 6 rows of the data.frame

```{r}
head(data_covid, 6)
```


### How many states exists?

(Hint: use `count()`)


```{r}
length(unique(data_covid$state))
# or
data_covid %>% group_by(state) %>% count()
```


## Data wrangling, Part 1

### Erase unnecessary rows

First remove the non-county(sic!) entries

```{r}
table(is.na(data_covid$county))
data_covid <- data_covid %>% filter(!is.na(county))
```


### Create a subset dataset

Find the latest date of the data, then subset the data with only the newest date for each country

```{r}
dat2 <- data_covid %>% filter(date == max(date))
```

### Max cases/deaths

Which county has the highest number of cases/deaths? What is the number?

```{r}
dat2 %>% group_by(county, state) %>% summarize(deaths = max(deaths)) %>% arrange(desc(deaths)) %>% head(5)

dat2 %>% group_by(county, state) %>% summarize(cases = max(cases)) %>% arrange(desc(cases)) %>% head(10)
```


## Data wrangling, Part 2


### First recorded cases/deaths (by county)

Find the 10 counties with the earliest recorded cases/deaths.

```{r}
data_covid %>% filter(cases > 0) %>% group_by(state, county) %>% arrange(desc(state, county, date)) %>% slice(1) %>% arrange(date) %>% select(county, state, date, cases) %>% head(10)

data_covid %>% filter(deaths > 0) %>% group_by(state, county) %>% arrange(desc(state, county, date)) %>% slice(1) %>% arrange(date) %>% select(county, state, date, deaths) %>% head(10)
```

### Aggregate the number of death

For each state, calculate the total number of cases/deaths for each day (Hint: grouping and summarize)


```{r}
data_covid %>% group_by(state, date) %>% summarise(cases = sum(cases),
                                                   deaths = sum(deaths))
```


### First recorded cases/deaths (by state)

Find the earliest day when a state recorded a case. Arrange by the date in descending order

```{r}
data_covid %>% filter(cases > 0) %>% group_by(state) %>% summarise(date = min(date)) %>% arrange(desc(date))
```

## Visualizing (Optional)

Now let's visualize the outputs using ggplot

### Time series plot of total cases/death in the US

```{r}
data_covid %>% group_by(date) %>% summarise(cases = sum(cases), deaths = sum(deaths)) %>% mutate(cases = cases / 1000, deaths = deaths / 1000) %>% pivot_longer(-date, names_to = "label", values_to = "count") %>% 
  ggplot(aes(x = date, y = count, group = label, color = label)) + geom_line()

data_covid %>% group_by(date) %>% summarise(cases = sum(cases)) %>% mutate(cases = cases / 1000) %>% ggplot(aes(x = date, y = cases)) + geom_line()

data_covid %>% group_by(date) %>% summarise(deaths = sum(deaths)) %>% mutate(deaths = deaths / 1000) %>% ggplot(aes(x = date, y = deaths)) + geom_line()

```


### Time series plot of total cases/deaths by the state

```{r}
data_covid %>% group_by(state, date) %>% summarise(cases = sum(cases)) %>% mutate(cases = cases / 1000) %>% ggplot(aes(x = date, y = cases)) + geom_line() + facet_wrap(~ state)

data_covid %>% group_by(state, date) %>% summarise(deaths = sum(deaths)) %>% mutate(deaths = deaths / 1000) %>% ggplot(aes(x = date, y = deaths)) + geom_line() + facet_wrap(~ state)
```


### Time series plot of total cases/deaths by the state per 100,000 inhabitants

```{r}
library(usa)
usapop <- usa::facts %>% select(state = name, population)
dat2 <- merge(x = data_covid, y = usapop, all.x = TRUE, by = "state")

dat2 %>% group_by(state, date, population) %>% summarise(cases = sum(cases)) %>% mutate(cases = cases / (population / 100000)) %>% ggplot(aes(x = date, y = cases)) + geom_line() + facet_wrap(~ state)

dat2 %>% group_by(state, date, population) %>% summarise(deaths = sum(deaths)) %>% mutate(deaths = deaths / (population / 100000)) %>% ggplot(aes(x = date, y = deaths)) + geom_line() + facet_wrap(~ state)
```

### Daily increase trend

How's the trend of the daily case increases. Which state look bad now?

Top 10 of states with most new cases in the last 7 days
```{r}
dat2 %>% group_by(state, date) %>% summarise(cases = sum(cases)) %>% mutate(trend = cases - lag(cases)) %>% filter(date >= max(date)-7) %>% group_by(state) %>% summarise(trend = sum(trend)) %>% arrange(desc(trend)) %>% head(10)
```

Top 10 of states with most new cases in the last 7 days in relation to population
```{r}
dat2 %>% group_by(state, population, date) %>% arrange(date) %>% summarise(cases = sum(cases)) %>% mutate(trend = cases - lag(cases)) %>% filter(date >= max(date)-7) %>% group_by(state, population) %>% summarise(increase_total = sum(trend)) %>% mutate(increase_per_10k = round(increase_total / (population / 10000), 1)) %>% arrange(desc(increase_per_10k)) %>% head(10)

```
